---
title: "DANL 210: Data Preparation and Management"
subtitle: Quarto Template<br><br> 
author: 
  - name: YOUR NAME
  - name: YOUR COWORKER 1
    
date: last-modified

execute:
  echo: true
  eval: true
  message: false
  warning: false
  fig-width: 9

format: 
  html:
    toc: true
    number-sections: true
    code-fold: false       # https://quarto.org/docs/output-formats/html-code.html
    code-tools: true
    code-link: true
    highlight-style: atom-one # atom-one tango espresso
---

```{r}
#| include: false

# install.packages("reticulate")
library(reticulate)

# Provide the pahname of the anaconda3 directory:
use_condaenv("C:/Users/mizzf/anaconda3/s", required = TRUE)
```


## Variables and Data Types

In Python, variables can store data of different types without explicitly declaring the type.

For example:

```{python}
print("Hello, world!")

```

## Control Structures

Python supports the usual logical conditions from mathematics:

```{python}
# Equals: a == b
# Not Equals: a != b
# Less than: a < b
# Less than or equal to: a <= b
# Greater than: a > b
# Greater than or equal to: a >= b
```

These conditions can be used in several ways, most commonly in 'if statements' and loops.

```{python}
# if statement:
if 5 > 2:
    print('Five is greater than two!')
```

## Functions

A function is a block of code which only runs when it is called.

You can pass data, known as parameters, into a function.

A function can return data as a result.


```{python}
# Defining a function:
def my_function():
    print('Hello from a function')

# Calling a function:
my_function()
```

## Lists and Dictionaries

A list is a collection which is ordered and changeable.

A dictionary is a collection which is unordered, changeable and indexed.

```{python}
# List example:
my_list = ['apple', 'banana', 'cherry']

# Dictionary example:
my_dict = {'name': 'John', 'age': 36}
```



# Pandas Basics - Orange Juice Sales

## Import Python libraries

```{python}
import pandas as pd
```



```{python}
oj = pd.read_csv('https://bcdanl.github.io/data/dominick_oj.csv')
```

```{python}
#| code-fold: true
#| code-summary: "Code!"

oj
```


```{python}
oj.describe()
```



## Python-R Interaction

Below is using Python's DataFrame `oj` to visualize using R's `ggplot`
```{r}
library(tidyverse)

# Access the Python pandas DataFrame
oj <- py$oj

# Plot using ggplot2
ggplot(oj, aes(x = log(sales), y = log(price), 
               color = brand)) +
  geom_point(alpha = .25) +
  geom_smooth(method = lm) +
  theme_minimal()
```

### Interactive DataFrame with R's `DT` Package

```{r}
#| echo: false
#| warning: false

# install.packages("DT")
library(DT)
datatable(oj)
```


In `*.ipynb` on Google Colab, we can use `itables` or just Google Colab's default to print DataFrame.

```{python}
#| eval: false

# !pip install itables
from itables import init_notebook_mode, show
init_notebook_mode(all_interactive=False)

oj = pd.read_csv('https://bcdanl.github.io/data/dominick_oj.csv')
show(oj)
```

